---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = F,
  warning = F,
  message = F,
  error = F
)
library(sf)
library(terra)
library(geodata)
library(tidyterra)
library(ggplot2)
library(dplyr)
library(SDMWorkflows)
library(rgbif)
```

# Klimaskovfond

<!-- badges: start -->
<!-- badges: end -->

The goal of Klimaskovfond is to generate the needed data sets to get analize the Klimaskovfond score system


```{r, cache=T}
Forest <- terra::rast("o:/Nat_BDR-data/Arealanalyse/2023/CLEAN/Rasterized/Rast_AllForerst_Croped.tif") |> as.numeric()

Desciduous <- terra::rast("o:/Nat_Ecoinformatics-tmp/JakobAssmann_au634851/treetype/treetype_bjer_dec.tif") 

Conifer <- terra::rast("o:/Nat_Ecoinformatics-tmp/JakobAssmann_au634851/treetype/treetype_bjer_con.tif") 


```
```{r asPolygons, cache=T}
Forest[Conifer== 1] <- NA

Forest_Poly <-  as.polygons(Forest) |> aggregate(by = "C_02") |> 
  disagg()
Forest_Poly$ID <- 1:nrow(Forest_Poly)
terra::writeVector(Forest_Poly, "Forest.shp", overwrite = T)
sf::write_sf(st_as_sf(Forest_Poly), "Forest.geojson")
```


```{r, echo = F, eval = F}
library(doParallel)
library(foreach)
require(doSNOW)
library(data.table)

cl <- makeSOCKcluster(50)

Forest_Poly <- terra::vect("Forest_Area.shp")

MoreThan2 <- Forest_Poly[Forest_Poly$Ha > 2,]

MoreThan2DF <- as.data.frame(MoreThan2)
MoreThan2DF$Richness <- NA

terra::writeVector(MoreThan2, "MoreThan2.shp")

doSNOW::registerDoSNOW(cl)

pb <- txtProgressBar(min=1, max=nrow(MoreThan2DF), style=3)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress=progress)
#v
Species <- foreach(i=1:nrow(MoreThan2DF), .packages = c("terra", "sf", "rgbif", "SDMWorkflows", "dplyr"), .options.snow=opts)%dopar% {
  try({
    MoreThan2 <- terra::vect("MoreThan2.shp")
    WKT <- MoreThan2[i,]|>
      terra::convHull() |>
      terra::project("epsg:4326") |>
      st_as_sf() |>
      st_as_sfc() |>
      st_as_text()

    Test <- rgbif::occ_count(hasCoordinate = T,
                             geometry = WKT,
                             year = '1999,2023',
                             facet="scientificName", facetLimit=10000)

    Temp <- SDMWorkflows::Clean_Taxa(Taxons = Test$scientificName) |>
      dplyr::select(kingdom, phylum, class, family, genus, species) |>
      dplyr::distinct()

    Temp$ID <- i
    MoreThan2DF[i,]$Richness <- nrow(Temp)
    message(paste(i, "of", nrow(MoreThan2DF) ,"ready", Sys.time()))
    Temp

  })
}

readr::write_csv(MoreThan2DF, "RichnessParallel.csv")
stopCluster(cl)

Species <- Species |> purrr::keep(is.data.frame) |> purrr::map(data.table::as.data.table) |> data.table::rbindlist()

saveRDS(Species, "Species.rds")

MoreThan2DF <- as.data.frame(MoreThan2) |> dplyr::select(Ha)
MoreThan2DF$ID <- 1:nrow(MoreThan2DF)

MoreThan2DF <- MoreThan2DF[MoreThan2DF$ID %in% unique(Species$ID),]

Species_Animals_Plants <- Species[kingdom %chin% c("Animalia", "Plantae")]

Species_Animals_Plants_N <- Species_Animals_Plants[, .N, by = ID]

Species_Animals_Plants_N <- Species_Animals_Plants_N |> 
  dplyr::left_join(MoreThan2DF)

write_csv(Species_Animals_Plants_N, "Animals_plants_richness.csv")

Species_Plants <- Species[kingdom %chin% c("Plantae")]

Species_Plants_N <- Species_Plants[, .N, by = ID]

Species_Plants_N <- Species_Plants_N |> 
  dplyr::left_join(MoreThan2DF)

write_csv(Species_Animals_Plants_N, "Plants_richness.csv")

Species_Animals <- Species[kingdom %chin% c("Animalia")]

Species_Animals_N <- Species_Animals[, .N, by = ID]

Species_Animals_N <- Species_Animals_N |> 
  dplyr::left_join(MoreThan2DF)

write_csv(Species_Animals_N, "Animals_richness.csv")
```


```{r HaArea, cache=T}
Forest_Poly <- terra::vect("Forest.shp")
Forest_Poly$Ha <- terra::expanse(Forest_Poly, unit= "ha") |> round(3)
terra::writeVector(Forest_Poly, "Forest_Area.shp", overwrite = T)
sf::write_sf(st_as_sf(Forest_Poly), "Forest_Area.geojson")

Forest_DF <- as.data.frame(Forest_Poly) |> dplyr::mutate(LogHa = log(Ha))
```


We have `r prettyNum(nrow(Forest_DF), big.mark = ",")` forest patches, whit are ranges from `r prettyNum(round(min(Forest_DF$Ha), 3), big.mark = ",")` to `r prettyNum(max(Forest_DF$Ha), big.mark = ",")` hectares, the distribution of areas can be seen here

```{r histogramplot}
ggplot(Forest_DF, aes(x = Ha)) + geom_histogram(bins = nclass.Sturges(Forest_DF$LogHa)) + scale_x_log10(labels = scales::comma) + scale_y_continuous(labels = scales::comma)
```


```{r, echo = F, cache=T}
Forest_Poly <- terra::vect("Forest_Area.shp")

MoreThan2 <- Forest_Poly[Forest_Poly$Ha > 2,]

MoreThan2DF <- as.data.frame(MoreThan2)

MoreThan2DF <- MoreThan2DF |> dplyr::mutate(LogHa = log(Ha))
```


however of those plots only `r prettyNum(nrow(MoreThan2DF), big.mark = ",")` of them correspond to plots with more than 2 Ha, we can see the distribution of those in the following plot

```{r histogramplot2}
ggplot(MoreThan2DF, aes(x = Ha)) + geom_histogram(bins = nclass.Sturges(Forest_DF$LogHa)) + scale_x_log10(labels = scales::comma) + scale_y_continuous(labels = scales::comma)
```

## Species area relationship

to find a better solution at the amount of Hectares necessary for a plot to 
